---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "40%"
)
```
# ERforResearch

This package implements the functionality to use the expected ranks (ER) in the context of research evaluation as presented in [this recent preprint](...). For a demonstration of the major functions check out the vignette (Demo ERforResearch). 

## Installation

The package can be downloaded from github using the `devtools`-package:

```{r, echo = TRUE, eval = FALSE}
# First, you might have to install devtools from CRAN
install.packages("devtools")

# Then install ERforResearch
devtools::install_github("snsf-data/ERforResearch")
```


## Requirement

The Bayesian Hierarchical Models for the ER are fitted in JAGS using `rjags`. Therefore you have to install the `rjags` from CRAN and JAGS following the instructions from [here](http://mcmc-jags.sourceforge.net/).
